Quantum Pseudo-Telepathy

### 2023-05-15 paper/main.tex
Step by step

A \emph{Magic Square} is a \(3 \times 3\) matrix whose entries are
\{-1, 1\} (or sometimes \{0,~1\}), with the property that the sum of
each row is \(1\) and the sum of each column is \(1\). The Magic
Square game is defined as follows: a referee Charlie asks Alice to
fill the entries of a row \(x \in \{0, 1, 2\}\), which he chooses
uniformly at random. Charlie, then, asks Bob to fill the entries of a
column \(y \in \{0, 1, 2\}\), also chosen at random. Alice and Bob win
if the product of the row given by Alice is \(1\), the product of the
column given by Bob is \(-1\) and the intersection of the given row
and column agrees. A successful implementation of a quantum winning
strategy for any pseudo-telepathy game must convince an observer that
something classicaly impossible is
happening\cite{brassard:2005}. Fortunately, as we shall see, proving
the classical impossibility of the Magic Square game is extremely easy.

We shall rely on the following references (\todo{add proper refs with
  bib})
\begin{itemize}
\item Quantum Pseudo-Telepathy (Gilles Brassard, Anne Broadbent and
  Alain Tapp; Magic Square = Sec.\ 5)
\end{itemize}

>>>>

A \emph{Magic Square} is a \(3 \times 3\) matrix whose entries are
\{-1, 1\} (or sometimes \{0,~1\}), with the property that the sum of
each row is \(1\) and the sum of each column is \(-1\). The Magic
Square game is defined as follows: a referee Charlie asks Alice to
fill the entries of a row \(x \in \{0, 1, 2\}\), which he chooses
uniformly at random. Charlie, then, asks Bob to fill the entries of a
column \(y \in \{0, 1, 2\}\), also chosen at random. Alice and Bob win
if the product of the row given by Alice is \(1\), the product of the
column given by Bob is \(-1\) and the intersection of the given row
and column agrees.

A successful implementation of a quantum winning strategy for any
pseudo-telepathy game must convince an observer that something
classicaly impossible is happening\cite{brassard:2005}. Fortunately,
proving the classical impossibility of the Magic Square game is
extremely easy. Assuming a deterministic strategy, even if Alice and
Bob are allowed to talk \emph{before} the game starts, the best that
they can do is to construct two grids in advance, one for Alice and
another for Bob, and give the entries according to the matrices
that they prepared. However, the product of Alice's matrix will always
be \(1\) and the product of Bob's matrix will always be \(-1\), so
there must be one cell that does not agree. Probabilistic strategies
cannot do better than deterministic~\cite[Theorem 2]{brassard:2005}, so
the Magic Square game cannot be won under classical assumptions.

We will approach this project as follows:

\begin{enumerate}
\item Describe the quantum Magic Square game in details;
\item Discuss classical solutions: illustrate with examples and
  implement them (and make the code available through a jupyter
  notebook).
\item Demonstrate that no classical strategy can win with probability
  1. Show that the best that a classical strategy can achieve is
  is win 8 out of 9 rounds;
\item Discuss the quantum solution, using entanglement: illustrate
  with examples;
\item Demonstrate that the quantum strategy can win with probability
  1; and
\item Implement the quantum solution in Qiskit
  (and make the code available through a jupyter notebook).
\end{enumerate}

We will rely on the following references (\todo{add proper refs with
  bib})
\begin{itemize}
\item Quantum Pseudo-Telepathy (Gilles Brassard, Anne Broadbent and
  Alain Tapp; Magic Square = Sec.\ 5)
\end{itemize}
